Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

### 2018-11-21 review/response.md

1. While this is an important and timing topic dealing with data uncertainty
   and privacy in the mobile age, the reviewers have raised concerns on
	- Clarity of the methodology
	- Lack of discussion of wifi-based human activity studies in the big data
	age
	- Proper use of terminology
	- Performance of the algorithm etc.

2. In lieu of these comments, they have recommended to make substantial
revisions to your manuscript

3. Also, since this is a "smart cities" related special issue, I ask you to
	- Add more discussions on smart cities in the introduction and conclusion
	sections and
	- Link your research more tightly to this theme.

> Thank you for the very helpful comments of the reviewers of our paper/ As
> requested, we have responded to all comments without adding significant
> length to the paper. We have also extended our discussion of how effective
> measurement of footfall is relevant to understanding the functioning of the
> Smart City. If any of what follows and what has been uploaded is unclear,
> please do not hesitate to contact me.
> 
> Yours sincerely,
>
> Balamurugan Soundararaj,
>
> (on behalf also of Paul Longley and James Cheshire)

### 2018-11-21 review/response.md

1. While this is an important and timing topic dealing with data uncertainty
   and privacy in the mobile age, the reviewers have raised concerns on
	- Clarity of the methodology
	- Lack of discussion of wifi-based human activity studies in the big data
	age
	- Proper use of terminology
	- Performance of the algorithm etc.

2. In lieu of these comments, they have recommended to make substantial
revisions to your manuscript

3. Also, since this is a "smart cities" related special issue, I ask you to
	- Add more discussions on smart cities in the introduction and conclusion
	sections and
	- Link your research more tightly to this theme.

> Thank you for the very helpful comments of the reviewers of our paper/ As
> requested, we have responded to all comments without adding significant
> length to the paper. We have also extended our discussion of how effective
> measurement of footfall is relevant to understanding the functioning of the
> Smart City. If any of what follows and what has been uploaded is unclear,
> please do not hesitate to contact me.
> 
> Yours sincerely,  
> Balamurugan Soundararaj,  
> (on behalf also of Paul Longley and James Cheshire)

@@

> Lugomer, K., & Longley, P. (2018). Towards a Comprehensive Temporal 
> Classification of Footfall Patterns in the Cities of Great Britain (Short
> Paper). In LIPIcs-Leibniz International Proceedings in Informatics (Vol. 114).
> Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik.

> There was no reason for
> selecting the time in the pilot study. the location was chosen for it it known
> to have high number of randomised probe requests and has one of the highest
> footfall in London. For the main study the counting was done such that each
> location has atleast 2/3 counts in the same time interval.

2. Will you give more detailed information about the data and the parameters for 
various clustering algorithms? Including the number of records, the number of 
clusters. How about the performance for different clustering methods 
in your study？

> Thank you. This part has been expanded and clarified in the revised manuscript. 
> The performance of the algorithms used for one dimensional clustering were 
> evaluated on the basis of the time taken for processing and the resulting 
> mean absolute percentage error. We found that the k-means gives
> us the best data for the quickest processing time. The graph based clustering
> was done using two parameters - sequence threshold of 60 and time threshold of
> 16 seconds. These values were arrived arrived at using trial and error by using
> a know device as the control mechanism. This has been illustrated further in 
> the revised manuscript.

### 2018-11-21 paper/sections/introduction.tex
Pre final

To this end, it is important to first undertake a thorough conceptual and
technical appraisal of our consumer data source.  In technical terms, screening
the information present in the `probes requests' and classifying them based on
their characteristics is essential in order to remove the ones emitted by
devices that do not indicate pedestrian activity, such as network enabled
printers and other fixed devices.  Related to this, a method to fingerprint
Wi-Fi probes is necessary to remove probes from individuals’ devices that in
conceptual terms should not be considered part of footfall – as when, for
example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first
individuals may carry multiple devices, or no device at all; and second, the
positioning and orientation of the sensor in the retail unit may lead to
systematic over- or under-enumeration.  These sources of bias in measurement
must be accommodated by manual recording of footfall at each location and
generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be
undertaken in parallel with technical profiling of the mix of consumer mobile
devices that probe our sensors, since the effectiveness of data cleaning
procedures discussed in this paper differ between individual locations and
configurations.

>>>>

To this end, it is important to first undertake a thorough conceptual and
technical appraisal of our consumer data source.  In technical terms, screening
the information present in the `probes requests' and clustering them based on
their characteristics is essential in order to remove the ones emitted by
devices that do not indicate pedestrian activity, such as network enabled
printers and other fixed devices.  Related to this, a method to fingerprint
Wi-Fi probes is necessary to remove probes from individuals’ devices that in
conceptual terms should not be considered part of footfall – as when, for
example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first
individuals may carry multiple devices, or no device at all; and second, the
positioning and orientation of the sensor in the retail unit may lead to
systematic over- or under-enumeration.  These sources of bias in measurement
must be accommodated by manual recording of footfall at each location and
generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be
undertaken in parallel with technical profiling of the mix of consumer mobile
devices that probe our sensors, since the effectiveness of data cleaning
procedures discussed in this paper differ between individual locations and
configurations.

### 2018-11-21 paper/sections/methodology.tex
Pre final

The nodes were then assigned a unique id based on the unique connected component
they belong to as shown in Figure \ref{methodology_schematic}. This unique
identifier is used in the place of MAC address for aggregation for the
anonymised probe requests. Though the recycling of sequence number after 4096
leads to multiple unique ids being reported from a single device, the magnitude
of error is greatly reduced.

>>>>

The nodes were then assigned a unique id based on the unique connected component
they belong to as shown in Figure \ref{methodology_schematic}. This unique
identifier is used in the place of MAC address for aggregation for the
anonymised probe requests. Though the recycling of sequence number after 4096
leads to multiple unique ids being reported from a single device from a sample
experiment done on randomised probe requests sent by "Google" devices we found
that the sequence number is reset for around 0.05\% of them. We hence assume
this to be inconsequential.

### 2018-11-21 paper/sections/pilot_study.tex
Pre final

To conclude, from the pilot study we found that both filtering and clustering
methods we devices work on complex real world data and results in a final
pedestrian counts within a MAPE of 20\%. We also found `k-means' and `quantile'
are best algorithms for classifying signal strengths and the threshold for time
and sequence numbers for the clustering algorithm is around 16 and 60
respectively.

>>>>

To conclude, from the pilot study we found that both filtering and clustering
methods we devices work on complex real world data and results in a final
pedestrian counts within a MAPE of 20\%. We also found `k-means' and `quantile'
are best algorithms for clustering signal strengths and the threshold for time
and sequence numbers for the clustering algorithm is around 16 and 60
respectively.
